\pkg{sparseHessianFD}: An R Package for Estimating Sparse Hessian Matrices

### 2016-12-24 nobuild/jss_revision/sparseHessianFD.tex
just about done with revision

\usepackage{Sweave}

@@

%%\newenvironment{knitrout}{\begin{figure}}{\end{figure}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

{\setlength{\extrarowheight}{.25em}
\begin{table}[ptb]
\begin{tabularx}{\linewidth}{>{\bfseries}r X}
  \toprule
x& A numeric vector, with length $M$ at which the object will be
  initialized and tested.\\
fn,gr& \proglang{R} Functions that return the value of the
  objective function, and its gradient. The first argument is the numeric
  variable vector.  Other named arguments can be passed to \func{fn}
  and {gr} as well (see the \funcarg{...} argument below).\\
rows, cols& Sparsity pattern: integer vectors of the row and column indices of
  the non-zero elements in the \emph{lower triangle} of the Hessian.\\
direct& This argument is deprecated, and is included only for
  backwards compatibility with earlier versions.\\
delta& The perturbation amount for finite differencing of the
  gradient to compute the Hessian (the $\delta$ in Section~\ref{sec:numdiff}).  Defaults to
  \code{sqrt(.Machine\$double.eps)}.\\
index1& If \variable{TRUE} (the default), \funcarg{row} and \funcarg{col} use one-based
  indexing.  If \variable{FALSE}, zero-based indexing is used.\\
  complex&If \variable{TRUE}, the complex step method is used.
                If \variable{FALSE} (the default), a simple finite
                differencing of gradients is used.\\
...& Additional arguments to be passed to \func{fn} and \func{gr}.\\
 \bottomrule
\end{tabularx}
\caption{Arguments to the sparseHessianFD initializer.}\label{tab:init}
\end{table}
}

@@

\begin{Schunk}
\begin{Sinput}
R> hs <- obj$hessian(P)
R> mean(abs(hs-true.hess))/mean(abs(hs))
\end{Sinput}
\begin{Soutput}
[1] 1.1217e-08
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> obj2 <- sparseHessianFD(P, fn=binary.f, gr=binary.grad,
+   rows=pattern[["rows"]], cols=pattern[["cols"]],
+   complex=TRUE,
+   data=binary, priors=priors, order.row=order.row)
R> hs2 <- obj2$hessian(P)
R> mean(abs(hs2-true.hess))/mean(abs(hs2))
\end{Sinput}
\begin{Soutput}
[1] 6.7543e-18
\end{Soutput}
\end{Schunk}

>>>>

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

%%\newenvironment{knitrout}{\begin{figure}}{\end{figure}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

{\setlength{\extrarowheight}{.25em}
\begin{table}[ptb]
\begin{tabularx}{\linewidth}{>{\bfseries}r X}
  \toprule
x& A numeric vector, with length $M$ at which the object will be
  initialized and tested.\\
fn,gr& \proglang{R} Functions that return the value of the
  objective function, and its gradient. The first argument is the numeric
  variable vector.  Other named arguments can be passed to \func{fn}
  and {gr} as well (see the \funcarg{...} argument below).\\
rows, cols& Sparsity pattern: integer vectors of the row and column indices of
  the non-zero elements in the \emph{lower triangle} of the Hessian.\\
delta& The perturbation amount for finite differencing of the
  gradient to compute the Hessian (the $\delta$ in Section~\ref{sec:numdiff}).  Defaults to
  \code{sqrt(.Machine\$double.eps)}.\\
index1& If \variable{TRUE} (the default), \funcarg{row} and \funcarg{col} use one-based
  indexing.  If \variable{FALSE}, zero-based indexing is used.\\
  complex&If \variable{TRUE}, the complex step method is used.
                If \variable{FALSE} (the default), a simple finite
                differencing of gradients is used.\\
...& Additional arguments to be passed to \func{fn} and \func{gr}.\\
 \bottomrule
\end{tabularx}
\caption{Arguments to the sparseHessianFD initializer.}\label{tab:init}
\end{table}
}

@@

\begin{Schunk}
\begin{Sinput}
R> hs <- obj$hessian(P)
R> mean(abs(hs-true.hess))/mean(abs(hs))
\end{Sinput}
\begin{Soutput}
[1] 2.3357e-09
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> obj2 <- sparseHessianFD(P, fn=binary.f, gr=binary.grad,
+   rows=pattern[["rows"]], cols=pattern[["cols"]],
+   complex=TRUE,
+   data=binary, priors=priors, order.row=order.row)
R> hs2 <- obj2$hessian(P)
R> mean(abs(hs2-true.hess))/mean(abs(hs2))
\end{Sinput}
\begin{Soutput}
[1] 8.0555e-17
\end{Soutput}
\end{Schunk}
